Model of lens candidates from Space Warps CFHTLS

### 2016-07-05 paper2.tex
editing

\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

A passing remark in the {\em Feynman Lectures}\footnote{
  \cite{1963flp..book.....F}} --- ``if you are looking for a good
problem, the exact details of how the arms are formed and what
determine the shapes of these galaxies has not been worked out'' ---
tells us that circa 1960, galaxy dynamics was expected be an easy
research puzzle, whose solution was just around the corner.  Yet a
half-century later, galaxies are not nearly as well-understood as
stars were then.  Two complications, which are clear now but were not
so clear then, are: (i)~dynamical chaos, because galaxies are not
spherical, and (ii)~dark matter.

The CFHTLS \citep{2012SPIE.8448E..0MC} has been a rich source of
lenses by arc-finders, by machine learning \citep{2016arXiv160504309P}
and by visual inspection \citep{2016MNRAS.455.1171M}.  The latter is
of particular interest here.

>>>>

\begin{keywords}
gravitational lensing: strong -- keyword2
\end{keywords}

@@

Light deflection at the rim of the Sun is famously $1.75''$.  The
deflection angle can also be expressed in terms of the escape velocity
$2v_{\rm esc}^2/c^2$, with $v_{\rm esc}$ at the solar surface being
$\simeq620\rm\,km/s$.  The Galactic escape velocity at the Sun's
location is similar: $\approx500\rm\,km/s$, typical of massive
galaxies.  Thus, the outer regions of massive galaxies have lensing
deflections comparable to that at the rim of the Sun.  Yet these
quantitatively similar deflections have qualitatively different
consequences.  Light deflection by the Sun must be accounted for in
modern astrometry \citep[see e.g.,][]{2015CQGra..32p5008C} but is not
itself a physics probe in the way it was a century ago.  With distant
galaxies, however, light bending by $\sim1''$ introduces two new
phenomena.  First, for galaxies distant enough that their apparent
size is comparable to the bending angle, light deflection can cause
multiple images of background sources, or strong lensing.  Second, the
gravitational field is dominated by dark matter, hence strong lensing
becomes a probe of dark matter in galaxies.

With no unambiguous dark-matter particle detections so far, dark
matter has studied only through its indirect consequences on galactic
and larger scales.  In most work, dark matter is taken to be a
collisionless non-relativistic fluid (cold dark matter or CDM); in
recent years, the simulation by \cite{2005Natur.435..629S} has been
particularly influential in studying the formation of CDM structures.
Other scenarios have also been considered, such as a cold condensing
boson fluid \citep{2016ApJ...818...89S}.  There is a general
consensus, however, that the origins of galaxy lie in the
gravitational collapse, fragmentation, and mergers of dark-matter
clumps, into which feel gas, cooling through radiative processes to
form dense clouds and eventually stars.  There is, however, much
debate about the details, of which \cite{2012RAA....12..917S} provide
a nice summary.

Strong-lensing galaxies are a useful source of information on the
mutual dynamics and dark matter and gas in galaxies.  The topic has
been explored in several studies
\citep{2009ApJ...703L..51K,2011ApJ...740...97L,2012MNRAS.424..104L,
  2016MNRAS.459.3677L,2016MNRAS.456..870B} but it is desirable to
enlarge the samples from tens of lensing galaxies to thousands.  Doing
so requires finding more lenses, of course, but also modelling their
masses.

Recent searches through the CFHTLS \citep{2012SPIE.8448E..0MC} using
arc-finders
\citep{2012ApJ...749...38M,2014A&A...567A.111M,2014ApJ...785..144G} by
machine learning \citep{2016arXiv160504309P} and by visual inspection
\citep{2016MNRAS.455.1191M} have between them discovered an average of
four lenses per square degree.  So one can be optimistic about finding
many thousands of lenses in the next generation of wide-field surveys.
But modelling them all will be very difficult if the process is
labour-intensive.

One possibility is to enlist the help of citizen-scientists for
modelling large samples of lenses.  \cite{2015MNRAS.447.2170K}
developed a system whereby experienced but non-professional lens
enthusiasts can model newly-discovered lenses collaboratively.
